---
title: "FinalProject"
author: "SHASHIDHAR REDDY MALIGIREDDY"
date: "2023-12-07"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
sal<- read.csv("salaries (1).csv")
```


```{r}
#null values
sum(is.na(sal))
```

```{r}
#unique values in each column
sapply(sal, function(x) length(unique(x)))
```


```{r}
#label encoding for `experience_level` column INTO `experience_level_label`
sal$experience_level_label <- as.numeric(factor(sal$experience_level))
```


```{r}
#label encoding for employment_type column INTO employment_type_label
sal$employment_type_label <- as.numeric(factor(sal$employment_type))
```



```{r}
#label encoding for company_size column INTO company_size_label
sal$company_size_label <- as.numeric(factor(sal$company_size))
```

```{r}
head(sal)
```

```{r}
library(readr)
library(ggplot2)
sal<- read_csv("salaries (1).csv")
head(sal)
```

```{r}
#if salary_currency is other than usd then convert salary_in_usd to USD
sal$salary_in_usd <- ifelse(sal$salary_currency == "AUD", sal$salary_in_usd * 0.65, sal$salary_in_usd)
sal$salary_in_usd <- ifelse(sal$salary_currency == "BRL", sal$salary_in_usd * 0.2, sal$salary_in_usd)
sal$salary_in_usd <- ifelse(sal$salary_currency == "CAD", sal$salary_in_usd * 0.74, sal$salary_in_usd)
sal$salary_in_usd <- ifelse(sal$salary_currency == "CHF", sal$salary_in_usd * 1.14, sal$salary_in_usd)
sal$salary_in_usd <- ifelse(sal$salary_currency == "DKK", sal$salary_in_usd * 0.14, sal$salary_in_usd)
sal$salary_in_usd <- ifelse(sal$salary_currency == "EUR", sal$salary_in_usd * 1.08, sal$salary_in_usd)
sal$salary_in_usd <- ifelse(sal$salary_currency == "GBP", sal$salary_in_usd * 1.26, sal$salary_in_usd)
sal$salary_in_usd <- ifelse(sal$salary_currency == "HKP", sal$salary_in_usd * 0.13, sal$salary_in_usd)
sal$salary_in_usd <- ifelse(sal$salary_currency == "HUF", sal$salary_in_usd * 0.0028, sal$salary_in_usd)
sal$salary_in_usd <- ifelse(sal$salary_currency == "ILS", sal$salary_in_usd * 0.27, sal$salary_in_usd)
sal$salary_in_usd <- ifelse(sal$salary_currency == "INR", sal$salary_in_usd * 0.012, sal$salary_in_usd)
sal$salary_in_usd <- ifelse(sal$salary_currency == "JPY", sal$salary_in_usd * 0.0068, sal$salary_in_usd)
sal$salary_in_usd <- ifelse(sal$salary_currency == "MXN", sal$salary_in_usd * 0.058, sal$salary_in_usd)
sal$salary_in_usd <- ifelse(sal$salary_currency == "NOK", sal$salary_in_usd * 0.091, sal$salary_in_usd)
sal$salary_in_usd <- ifelse(sal$salary_currency == "PHP", sal$salary_in_usd * 0.018, sal$salary_in_usd)
sal$salary_in_usd <- ifelse(sal$salary_currency == "PLN", sal$salary_in_usd * 0.25, sal$salary_in_usd)
sal$salary_in_usd <- ifelse(sal$salary_currency == "SGD", sal$salary_in_usd * 0.75, sal$salary_in_usd)
sal$salary_in_usd <- ifelse(sal$salary_currency == "THB", sal$salary_in_usd * 0.028, sal$salary_in_usd)
sal$salary_in_usd <- ifelse(sal$salary_currency == "TRY", sal$salary_in_usd * 0.035, sal$salary_in_usd)
sal$salary_in_usd <- ifelse(sal$salary_currency == "ZAR", sal$salary_in_usd * 0.053, sal$salary_in_usd)
```


```{r}
#remove salary_currency column
#sal$salary_currency <- NULL
#sal$salary_in_usd<-salaries$salary_in_usd / 1000
```

```{r}
#access data with job_title as Data Scientist,AI Engineer,ML Engineer
salaries_1 <- subset(sal, job_title %in% c("Data Scientist", "AI Engineer", "ML Engineer"))
```

```{r}
unique(salaries_1$job_title)
```


#How does the salary distribution on vary across different experience levels and job Ntles within the AI, ML, and Data Science domains globally?

```{r}
ggplot(salaries_1, aes(x = experience_level, y = salary_in_usd, fill = job_title)) + geom_bar(stat = "identity", position = "dodge") + theme(axis.text.x = element_text(angle = 
90, hjust = 1))

```









```{r}

summary(sal)

cor_matrix <- cor(sal[, c("salary_in_usd", "remote_ratio")])
print("Correlation Matrix:")
print(cor_matrix)


unique_levels <- unique(sal$experience_level)


anova_result <- aov(salary_in_usd ~ experience_level, data = sal)


print(summary(anova_result))

posthoc_result <- TukeyHSD(anova_result)
print(posthoc_result)
```

here we started doing the dataanalysis for the dataset for the understanding purpose




```{r}
ggplot(sal, aes(x = salary_in_usd)) +
  geom_histogram(bins = 30, fill = "blue", color = "black") +
  theme_minimal() +
  labs(title = "Distribution of Salaries", x = "Salary in USD", y = "Frequency") +
  scale_x_continuous(
    labels = scales::number_format(scale = 1e-3, suffix = "k"),
    breaks = seq(0, max(sal$salary_in_usd), by = 50000)
  )
```
here we can see the distribution os salaries which are starting from 30k to 300k 


Is there a significant difference in salaries based on the extent of
remote work, and does this vary across countries or company sizes?

```{r}
library(dplyr)

sal %>%
  group_by(company_size) %>%
  summarise(avg_salary = mean(salary_in_usd)) %>%
  ggplot(aes(x = company_size, y = avg_salary, fill = company_size)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  theme_minimal() +
  labs(
    title = "Average Salary by Company Size",
    x = "Company Size",
    y = "Average Salary in USD"
  ) +
  scale_fill_manual(values = c("blue", "green", "red"))  # Custom colors



```

here we started grouping the salaries by company size and avaraged for the better understanding pourpose by the compoany size for the better representation of the graphs.



Are  experience levels consistently associated with
higher salaries across different company sizes?
```{r}
library(ggplot2)


sal$salary_in_k <- sal$salary_in_usd / 1000

ggplot(sal, aes(x = experience_level, y = salary_in_k, color = experience_level)) +
  geom_point() +
  theme_minimal() +
  labs(
    title = "Scatter Plot of Salary in 'k' by Experience Level",
    x = "Experience Level",
    y = "Salary in 'k'"
  ) +
  scale_y_continuous(breaks = seq(0, max(sal$salary_in_k), by = 100),
                     labels = scales::number_format(scale = 1000e-3)) +
  scale_color_brewer(palette = "Set3")  

```

from the above graph we can see that the salaries of each employe based on there experience level

```{r}

sample_size <- length(sal$salary_in_usd)
cat("Sample Size:", sample_size, "\n")
library(dplyr)

summary_data <- sal %>%
  group_by(company_size) %>%
  summarise(avg_salary = mean(salary_in_usd))

str(summary_data)
```

here we see that the avarage salary of large , medium and small companies salaries of employeses

```{r}

ggplot(sal, aes(x = as.factor(remote_ratio), y = salary_in_usd)) +
  geom_violin(fill = "skyblue", color = "blue") +
  labs(title = "Salary Distribution by Remote Work Ratio", x = "Remote Work Ratio", y = "Salary (USD)")

ggplot(sal, aes(x = as.factor(company_size), y = salary_in_usd)) +
  geom_violin(fill = "lightcoral", color = "red") +
  labs(title = "Salary Distribution by Company Size", x = "Company Size", y = "Salary (USD)")
```

```{r}
ggplot(sal, aes(x = as.factor(remote_ratio), y = salary_in_usd / 1000, color = as.factor(remote_ratio))) +
  geom_jitter(position = position_jitter(width = 0.3)) +
  labs(title = "Salary Differences by Remote Work Ratio", x = "Remote Work Ratio", y = "Salary (in 50ks USD)") +
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3, suffix = "ks")) +  # Adjust y-axis labels
  theme_minimal()

ggplot(sal, aes(x = as.factor(company_size), y = salary_in_usd / 1000, color = as.factor(company_size))) +
  geom_jitter(position = position_jitter(width = 0.3)) +
  labs(title = "Salary Differences by Company Size", x = "Company Size", y = "Salary (in 50ks USD)") +
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3, suffix = "ks")) +  # Adjust y-axis labels
  theme_minimal()
```


How does the salary distribuNon vary across different experience
levels and the company size?
```{r}
sal %>%
  group_by(experience_level, company_size, employee_residence) %>%
  summarise(avg_salary = mean(salary_in_usd)) %>%
  ggplot(aes(x = company_size, y = avg_salary, fill = experience_level)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Average Salaries by Experience Level, Company Size, and Region",
       x = "Company Size", y = "Average Salary (USD)", fill = "Experience Level") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

from the above graph we can see the relation between the avarage salary , company size and the experience level 


```{r}
salary_ranges <- c(0, 50000, 100000, 150000, 200000, Inf)
salary_labels <- c("Low", "Medium-Low", "Medium", "Medium-High", "High")

sal$salary_category <- cut(sal$salary_in_usd, breaks = salary_ranges, labels = salary_labels, include.lowest = TRUE)

head(sal)
```

 here we have grouped the salaries into four groups like low , medium-low ,medium , medium-high
 
 
 
 
 
 
how can we predict the salary of Machine Learning Engineers in the United States using features such as work year, experience level, and company size? 
 
```{r}

sal <- sal[, c( "experience_level", "company_size", "salary_in_usd")]
sal$experience_level <- as.numeric(as.factor(sal$experience_level))
sal$company_size <- as.numeric(as.factor(sal$company_size))

set.seed(123)
train_model <- sample(1:nrow(sal), 0.8 * nrow(sal))
train_data <- sal[train_model, ]
test_data <- sal[-train_model, ]

# Linear regression
lm_model <- lm(salary_in_usd ~   experience_level + company_size, data = train_data)
lm_predictions <- predict(lm_model, newdata = test_data)

# Random forest
library(randomForest)
rf_model <- randomForest(salary_in_usd ~ experience_level + company_size, data = train_data)
rf_predictions <- predict(rf_model, newdata = test_data)

# Evaluate models
lm_rmse <- sqrt(mean((lm_predictions - test_data$salary_in_usd)^2))
rf_rmse <- sqrt(mean((rf_predictions - test_data$salary_in_usd)^2))

# Evaluate the model on the test set
predictions <- predict(lm_model, newdata = test_data)

# R-squared
r_squared <- 1 - sum((test_data$salary_in_usd - predictions)^2) / sum((test_data$salary_in_usd - mean(test_data$salary_in_usd))^2)

# Mean Squared Error (MSE)
mse <- mean((test_data$salary_in_usd - predictions)^2)

cat("R-squared:", round(r_squared, 4), "\n")
cat("Mean Squared Error:", round(mse, 2), "\n")


cat("Linear Regression RMSE:", round(lm_rmse, 4), "\n")
cat("Random Forest RMSE:", round(rf_rmse, 4), "\n")
```
\
```{r}
#missing_values <- colSums(is.na(sal[c('experience_level', 'salary_category')]))

# Print the summary
#print(missing_values)

# Handle missing values (for example, removing rows with missing values)
sal <- na.omit(sal)

# Select relevant features
features <- sal %>%
  select(experience_level)

# Remove rows with missing values
#data <- na.omit(cbind(features, salary_category = sal$salary_category))
```










how can we predict the company level  based on their experience level and the salary ?



```{r}
library(randomForest)
library(Metrics)




# Convert factors to numeric
sal$experience_level <- as.factor(sal$experience_level)
sal$company_size <- as.factor(sal$company_size)

# Split the dataset

train_model <- createDataPartition(sal$company_size, p = 0.80, list = FALSE)
training_set <- sal[train_model,]
test_set <- sal[-train_model,]

# Train a Random Forest model
model <- randomForest(company_size ~ experience_level + salary, data = training_set)

# Make predictions
predictions <- predict(model, newdata = test_set)

# Evaluate the model
conf_matrix <- confusionMatrix(predictions, test_set$company_size)

# Calculate metrics
accuracy <- sum(predictions == test_set$company_size) / length(test_set$company_size)

# Output the results
list(confusion_matrix = conf_matrix, accuracy = accuracy)

numeric_predictions <- as.numeric(predictions)
numeric_test_labels <- as.numeric(test_set$company_size)

# Calculate MSE
mse <- mse(numeric_test_labels, numeric_predictions)

# Calculate RSE (Root Mean Squared Error)
rse <- sqrt(mse)

# Calculate R²
r_squared <- R2(numeric_test_labels, numeric_predictions)

# Output the results
list(mse = mse, rse = rse, r_squared = r_squared)

```

```{r}
# Load necessary libraries
library(tidyverse)
library(caret)
library(nnet)

# Convert categorical variables to factors
data_frame$experience_level <- as.factor(data_frame$experience_level)
data_frame$company_size <- as.factor(data_frame$company_size)

# Split the data into training and testing sets

splitIndex <- createDataPartition(data_frame$company_size, p = .80, list = FALSE)
training_set <- data_frame[splitIndex,]
test_set <- data_frame[-splitIndex,]

# Train logistic regression model

multinom_model <- multinom(company_size ~ experience_level + salary_in_usd, data = training_set)

# Make predictions
predictions <- predict(multinom_model, test_set)

# Evaluate the model
confusionMatrix(predictions, test_set$company_size)


```



```{r}

library(tidyverse)
library(caret)
library(e1071)

sal$experience_level <- as.factor(sal$experience_level)
sal$company_size <- as.factor(sal$company_size)

set.seed(123)  # for reproducible results
splitIndex <- createDataPartition(sal$company_size, p = .80, list = FALSE)
training_set <- sal[splitIndex,]
test_set <- sal[-splitIndex,]

# Train an SVM model
svm_model <- svm(company_size ~ experience_level + salary_in_usd, data = training_set)

# Make predictions
predictions <- predict(svm_model, test_set)

# Evaluate the model
confusionMatrix(predictions, test_set$company_size)
```

```{r}
colnames(sal)
```

how can we predict the salary of Machine Learning Engineers in the United States using features such as work year, experience level, and company size? 

```{r}
library(dplyr)
library(fastDummies)
library(glmnet)


features <- sal[c("work_year", "experience_level", "employment_type", "job_title", 
                  "salary", "salary_currency", "salary_in_usd", "employee_residence", 
                  "remote_ratio", "company_location", "company_size")]


features_encoded <- fastDummies::dummy_cols(features, select_columns = c("experience_level", "job_title",  "company_size"))

# Combine encoded features with the target variable (salary)
data <- cbind(features_encoded, salary_in_usd = sal$salary_in_usd)

# splitting the dataset into train and test data , 80% is train data and 20% is test data
set.seed(123)
train_split <- sample(seq_len(nrow(data)), size = 0.8 * nrow(data))
train_data <- data[train_split, ]
test_data <- data[-train_split, ]

# parameter tuning using cross-validation for the enhancement
cv_model <- cv.glmnet(
  x = as.matrix(train_data[, -ncol(train_data)]),
  y = train_data$salary_in_usd,
  alpha = 1  
)

# Find the optimal lambda for using in the lasso regression
best_lambda <- cv_model$lambda.min



# Build a final model using the optimal lambda
final_model <- glmnet(
  x = as.matrix(train_data[, -ncol(train_data)]),
  y = train_data$salary_in_usd,
  alpha = 1,  
  lambda = best_lambda
)

# Make predictions on the test set
predictions <- predict(final_model, newx = as.matrix(test_data[, -ncol(test_data)]))

# Evaluate the final model
mse <- mean((predictions - test_data$salary_in_usd)^2)
rsquared <- 1 - mse / var(test_data$salary_in_usd)

cat("Mean Squared Error:", mse, "\n")
cat("R-squared:", rsquared, "\n")
```


